\emph{Light Exposure Behavior Assessment (LEBA)}: Development of a novel instrument to capture light exposure-related behaviours

### 2022-01-30 manuscript.Rmd
Time-zone tibble is not working

  An expert panel prepared the initial 48 item pool. Responses to these items were then collected in an online survey producing responses from an international sample (690 completed responses, 74 countries, 28 time zones). Exploratory factor analysis (EFA) on an initial subset of our sample (n=428) rendered a five-factor solution with 25 items (Wearing blue light filters, spending time outdoors, using phone and smart-watch in bed, using light before bedtime, using light in the morning and during daytime). In a confirmatory factor analysis (CFA) performed on an independent subset of participants (n=262), we removed two further items to attain the best fit for the five-factor solution (CFI=0.97, TLI=0.96, RMSEA=0.05, SRMR=0.09). The internal consistency reliability coefficient for the total instrument was McDonald’s omega_total=0.73. Measurement model invariance analysis between native and non-native English speakers showed our model attained the highest level of invariance (residual invariance; CFI=0.95, TLI =0.95, RMSEA=0.05). Lastly, a short form of LEBA (n=18) was developed using Item Response Theory on the complete sample (n=690).

@@

To ensure construct adequacy we thoroughly assessed the current status of literature and identified a variety of light exposure related scales. However, no scales specifically measuring the behavioural component of light exposure were found (see Table \@ref(tab:eltab)). Consequentially we pursued to introduce a new openly available scale to address this research gap. For this purpose, an expert researcher panel from the fields of chronobiology, light research, neuroscience, and psychology – comprising all authors of the present manuscript – generated and collected preliminary item ideas. Special attention was paid to design items circumscribed to assess light exposure *behaviour* as opposed to subjective measurements of the light environment [@eklund1996development; @dianat2013objective]) and semi-quantitative assessments of light sources' illuminance [@bajaj2011validation] in order to maintain content validity. In a collective effort the generated items were then peer-reviewed, amended, unified, and complemented with a suitable response scale (5-point Likert-scale ranging from 1 "Never/Does not apply/I don’t know" to 5 "Always"). This process was finalized when all experts were in agreement, resulting in 48 items to implement in the data collection.

@@

Finally, we assessed age, sex, gender identity, occupational status, COVID-19 related occupational setting, time zone/country of residence and native language as single-item demographic variables. The demographic characteristics of our sample are given in Table \@ref(tab:demographics).

@@

# create descriptive table for the demographic vars (excluding tz & Country) with gtsummary
desctable.data %>%
  tbl_summary(
    by = IncludedInCFA,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 2,
    label = list(slypos_demographics_age ~ "Age",
                 slypos_demographics_sex.factor ~ "Sex",
                 slypos_demographics_gender.factor ~ "Gender-Variant Identity",
                 slypos_demographics_language.factor ~ "Native English Speaker",
                 slypos_demographics_work_or_school.factor ~ "Occupational Status",
                 slypos_demographics_school.factor ~ "Occupational setting"
                 ),
    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% add_p() %>% add_q() %>%
modify_header(label ~ "**Variable**") %>% separate_p_footnotes() %>%
  modify_caption("Demographic Characteristics") -> desc_table

@@

#summarise country/time zone data

descriptives.data%>%
  tbl_summary(
    label = slypos_demographics_tz.factor ~ "Time zone - Country",
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    missing = "no",
    sort = all_categorical() ~ "frequency",
    include = "slypos_demographics_tz.factor"
  )  %>% bold_labels() %>% 
  modify_header(label ~ "") %>%
  as_tibble(format = 'pipe') -> tz_tibble

@@

```{r demographics, warning=F, message=F}
desc_kable  %>% kable_styling(latex_options = c("scale_down")) %>% landscape()
```

Table \@ref(tab:demographics) summarizes the survey participants' demographic characteristics. Only participants completing the full LEBA questionnaire were included, thus there are no missing values in the item analyses. XX participants were excluded from analysis due to not passing at least one of the "attention check" items. For exploring initial factor structure (EFA), a sample of 250-300 is recommended [@comreyFirstCourseFactor1992; @schonbrodtWhatSampleSize2013]. For estimating the sample size for the confirmatory factor analysis (CFA) we followed the N:q rule [@bentlerPracticalIssuesStructural1987; @jacksonRevisitingSampleSize2003; @klinePrinciplesPracticeStructural2015; @worthingtonScaleDevelopmentResearch2006], where ten participants per parameter is required to earn trustworthiness of the result. Our sample size exceeds these requirements: Anonymous responses from a total of *n* = `r nrow(desctable.data)` participants were included in the analysis of the current study, split into samples for exploratory (EFA: *n* = `r nrow(desctable.data[desctable.data$IncludedInCFA == "1. EFA Sample",])`) and confirmatory factor analysis (CFA: *n* = `r nrow(desctable.data[desctable.data$IncludedInCFA == "2. CFA Sample",])`). The EFA sample included participants filling out the questionnaire from 17 May 2021 to XX XXX 2021, whereas participants who filled out the questionnaire from YY YYY 2021 to 3 September 2021 were included in the CFA analysis. Participants indicated filling out the online survey from a diverse range of geographic locations. The ten most common country and time zone combinations included:

- `r tz_tibble[2,1]`:     `r tz_tibble[2,2]` 
- `r tz_tibble[3,1]`:     `r tz_tibble[3,2]`
- `r tz_tibble[4,1]`:     `r tz_tibble[4,2]`
- `r tz_tibble[5,1]`:     `r tz_tibble[5,2]`
- `r tz_tibble[6,1]`:     `r tz_tibble[6,2]`
- `r tz_tibble[7,1]`:     `r tz_tibble[7,2]`
- `r tz_tibble[8,1]`:     `r tz_tibble[8,2]`
- `r tz_tibble[9,1]`:     `r tz_tibble[9,2]`
- `r tz_tibble[10,1]`:    `r tz_tibble[10,2]`
- `r tz_tibble[11,1]`:    `r tz_tibble[11,2]`

For a complete list of geographic locations, see Suppl. Table \@ref(tab:tztabprint).

@@

@@

## Developing Short form of LEBA
We sought the Item Response Theory (IRT) to develop the short form of LEBA. IRT the conventional classical test theory-based analysis by gathering information on item quality by indices like item difficulty, item discrimination, and item information [@bakerBasicsItemResponse2017]. Item is  judged based on item information in relation to participants' latent trait level ($\theta$). We fitted each factor of LEBA with the graded response model [@samejima1997handbook] to the combined EFA and CFA sample (n =690). Item discrimination indicates the pattern of variation in the categorical responses with the changes in latent trait level ($\theta$), and item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported the item discrimination parameter and only discarded the items with relatively flat item information curve (information <.2) to develop the short form of LEBA. @bakerBasicsItemResponse2017 categorized the item discrimination in as none = 0; very low =0.01 to 0.34; low = 0.35 to 0.64; moderate = 0.65 to 1.34 ; high = 1.35 to 1.69; very high >1.70. Table \@ref(tab:IRTparameters) summarizes the IRT parameters of our tool. Item discrimination parameters of our tool fell in very high (10 items), high (4 items), moderate (4 items), and low ( 5 items) categorizes indicating a good range of discrimination along the latent trait level ($\theta$). Examination of the item information curve (Sup.fig \@ref(fig:F2IF)-\@ref(fig:F5IF)) indicated 5 items (1, 25,  38, 30, & 41) had relatively flat information curves (I($\theta$) <.20) thus discarded creating a short form of LEBA with 5 factors and 18 items.

@@

ggsave("Manuscript.figures/allitemplots.png",allitemplots, width = 12, height = 12, dpi = 1200, bg = "white")

@@

TIC.plot <- cowplot::plot_grid(F1.tic,F2.tic, F3.tic,
                                   F4.tic,F5.tic,
                   labels = "AUTO",
                     align="v",
                     ncol = 2,
                     label_size = 15,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")
ggsave("Manuscript.figures/TIC.png",TIC.plot, width = 10, height = 12, dpi = 600, bg = "white")

>>>>

  An expert panel prepared the initial 48 item pool. Responses, consisting rating the frequency of engaging in the per-item behavior on a 5-point Likert type scale were  collected in an online survey producing responses from an international sample (690 completed responses, 74 countries, 28 time zones). Exploratory factor analysis (EFA) on an initial subset of our sample (n=428) rendered a five-factor solution with 25 items (Wearing blue light filters, spending time outdoors, using phone and smart-watch in bed, using light before bedtime, using light in the morning and during daytime). In a confirmatory factor analysis (CFA) performed on an independent subset of participants (n=262), we removed two further items to attain the best fit for the five-factor solution (CFI=0.97, TLI=0.96, RMSEA=0.05, SRMR=0.09). The internal consistency reliability coefficient for the total instrument was McDonald’s omega_total=0.73. Measurement model invariance analysis between native and non-native English speakers showed our model attained the highest level of invariance (residual invariance; CFI=0.95, TLI =0.95, RMSEA=0.05). Lastly, a short form of LEBA (n=18) was developed using Item Response Theory on the complete sample (n=690).

@@

To ensure construct adequacy we thoroughly assessed the current status of literature and identified a variety of light exposure related scales. However, no scales specifically measuring the behavioural component of light exposure were found (see Table \@ref(tab:reltab2)). Consequentially we pursued to introduce a new openly available scale to address this research gap. For this purpose, an expert researcher panel from the fields of chronobiology, light research, neuroscience, and psychology – comprising all authors of the present manuscript – generated and collected preliminary item ideas. Special attention was paid to design items circumscribed to assess light exposure *behaviour* as opposed to subjective measurements of the light environment [@eklund1996development; @dianat2013objective]) and semi-quantitative assessments of light sources' illuminance [@bajaj2011validation] in order to maintain content validity. In a collective effort the generated items were then peer-reviewed, amended, unified, and complemented with a suitable response scale (5-point Likert-scale ranging from 1 "Never/Does not apply/I don’t know" to 5 "Always"). This process was finalized when all experts were in agreement, resulting in 48 items to implement in the data collection.

@@

Finally, we assessed age, sex, gender identity, occupational status, COVID-19 related occupational setting, time zone/country of residence and native language as single-item demographic variables. The demographic characteristics of our sample are given in Table \@ref(tab:demotab).

@@

# create descriptive table for the demographic vars (excluding tz & Country) with gtsummary
desctable.data %>%
  tbl_summary(
    by = IncludedInCFA,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 2,
    label = list(slypos_demographics_age ~ "Age",
                 slypos_demographics_sex.factor ~ "Sex",
                 slypos_demographics_gender.factor ~ "Gender-Variant Identity",
                 slypos_demographics_language.factor ~ "Native English Speaker",
                 slypos_demographics_work_or_school.factor ~ "Occupational Status",
                 slypos_demographics_school.factor ~ "Occupational setting"
                 ),

    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% add_p() %>% add_q() %>%
modify_header(label ~ "**Variable**") %>% separate_p_footnotes() %>%
  modify_caption("Demographic Characteristics") -> desc_table

@@

#summarise country/time zone data
#The following chunk is not working now. May be package problem (Sorting command is not working)

# descriptives.data%>%
#   tbl_summary(
#     label = slypos_demographics_tz.factor ~ "Time zone - Country",
#     statistic = list(all_categorical() ~ "{n} ({p}%)"),
#     missing = "no",
#     sort = all_categorical() ~ "frequency",
#     include = "slypos_demographics_tz.factor"
#   )  %>% bold_labels() %>% 
#   modify_header(label ~ "") %>%
#   as_tibble(format = 'pipe') -> tz_tibble

@@

```{r demotab, warning=F, message=F}
desc_kable  %>% kable_styling(latex_options = c("scale_down")) %>% landscape()
```

Table \@ref(tab:demotab) summarizes the survey participants' demographic characteristics. Only participants completing the full LEBA questionnaire were included, thus there are no missing values in the item analyses. XX participants were excluded from analysis due to not passing at least one of the "attention check" items. For exploring initial factor structure (EFA), a sample of 250-300 is recommended [@comreyFirstCourseFactor1992; @schonbrodtWhatSampleSize2013]. For estimating the sample size for the confirmatory factor analysis (CFA) we followed the N:q rule [@bentlerPracticalIssuesStructural1987; @jacksonRevisitingSampleSize2003; @klinePrinciplesPracticeStructural2015; @worthingtonScaleDevelopmentResearch2006], where ten participants per parameter is required to earn trustworthiness of the result. Our sample size exceeds these requirements: Anonymous responses from a total of *n* = `r nrow(desctable.data)` participants were included in the analysis of the current study, split into samples for exploratory (EFA: *n* = `r nrow(desctable.data[desctable.data$IncludedInCFA == "1. EFA Sample",])`) and confirmatory factor analysis (CFA: *n* = `r nrow(desctable.data[desctable.data$IncludedInCFA == "2. CFA Sample",])`). The EFA sample included participants filling out the questionnaire from 17 May 2021 to XX XXX 2021, whereas participants who filled out the questionnaire from YY YYY 2021 to 3 September 2021 were included in the CFA analysis. Participants indicated filling out the online survey from a diverse range of geographic locations. The ten most common country and time zone combinations included:

<!-- - `r tz_tibble[2,1]`:     `r tz_tibble[2,2]` -->
<!-- - `r tz_tibble[3,1]`:     `r tz_tibble[3,2]` -->
<!-- - `r tz_tibble[4,1]`:     `r tz_tibble[4,2]` -->
<!-- - `r tz_tibble[5,1]`:     `r tz_tibble[5,2]` -->
<!-- - `r tz_tibble[6,1]`:     `r tz_tibble[6,2]` -->
<!-- - `r tz_tibble[7,1]`:     `r tz_tibble[7,2]` -->
<!-- - `r tz_tibble[8,1]`:     `r tz_tibble[8,2]` -->
<!-- - `r tz_tibble[9,1]`:     `r tz_tibble[9,2]` -->
<!-- - `r tz_tibble[10,1]`:    `r tz_tibble[10,2]` -->
<!-- - `r tz_tibble[11,1]`:    `r tz_tibble[11,2]` -->

For a complete list of geographic locations, see Suppl. Table \@ref(tab:tztable).

@@

```{r factorsplot, eval=FALSE, include=FALSE}
#This chunk holds code for compiled factor identification plots

fator2 <- cowplot::plot_grid(parallel, scree.plot,Hull, 
                   labels = "AUTO",
                   ncol=3, 
                  align = "v", 
                  label_fontfamily = "Arial",
                  label_fontface = "plain")
ggsave("Manuscript.figures/factors2.png",fator2, width = 18, height = 6, dpi = 600)
```

@@

## Developing Short form of LEBA
We sought the Item Response Theory (IRT) to develop the short form of LEBA. IRT the conventional classical test theory-based analysis by gathering information on item quality by indices like item difficulty, item discrimination, and item information [@bakerBasicsItemResponse2017]. Item is  judged based on item information in relation to participants' latent trait level ($\theta$). We fitted each factor of LEBA with the graded response model [@samejima1997handbook] to the combined EFA and CFA sample (n =690). Item discrimination indicates the pattern of variation in the categorical responses with the changes in latent trait level ($\theta$), and item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported the item discrimination parameter and only discarded the items with relatively flat item information curve (information <.2) to develop the short form of LEBA. @bakerBasicsItemResponse2017 categorized the item discrimination in as none = 0; very low =0.01 to 0.34; low = 0.35 to 0.64; moderate = 0.65 to 1.34 ; high = 1.35 to 1.69; very high >1.70. Table \@ref(tab:IRTparameters) summarizes the IRT parameters of our tool. Item discrimination parameters of our tool fell in very high (10 items), high (4 items), moderate (4 items), and low ( 5 items) categorizes indicating a good range of discrimination along the latent trait level ($\theta$). Examination of the item information curve (Sup.fig \@ref(fig:IF)) indicated 5 items (1, 25,  38, 30, & 41) had relatively flat information curves (I($\theta$) <.20) thus discarded creating a short form of LEBA with 5 factors and 18 items.

@@

ggsave("Manuscript.figures/allitemplots.png",allitemplots, width = 18, height = 12, dpi = 600, bg = "white")

discarded.items.plot <- cowplot::plot_grid(F3.iteminfo.4,
                                           F3.iteminfo.5,
                                           F4.iteminfo.3,
                                           F5.iteminfo.3,
                                           F5.iteminfo.5, 
                                            labels = "AUTO",
                                   align="v",
                                   ncol = 5,
                                   label_size = 12,
                                   label_fontfamily = "Arial",
                                   label_fontface = "plain")

ggsave("Manuscript.figures/discarded.png",discarded.items.plot, width = 18, height = 4, dpi = 600, bg = "white")

@@

TIC.plot <- cowplot::plot_grid(F1.tic,F2.tic, F3.tic,
                                   F4.tic,F5.tic,
                   labels = "AUTO",
                     align="v",
                     ncol = 3,
                     label_size = 15,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")
ggsave("Manuscript.figures/TIC.png",TIC.plot, width = 18, height = 12, dpi = 600, bg = "white")

### 2022-01-30 s1.Rmd
Time-zone tibble is not working

mainfont: Arial
---
```{r library, include=FALSE}

@@

```{r setup, include = FALSE}
library("papaja")
r_refs("references.bib")
```

@@

>>>>

mainfont: Arial
---
```{r setup, warning=FALSE, include=FALSE}
#This chunk controls the common chunk parameters for the whole manuscript
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, dpi=600, fig.width = 6,
 fig.asp = 0.8,out.width = "80%", dev=c('png','postscript'))
options(knitr.duplicate.label = "allow")
set.seed(123)
par(family = "Arial")

```{r library1, include=FALSE}

@@

@@

```{r library, include=FALSE}
#This chunk holds code for calling the required packages

library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(car)#reversecoding 
library(psych)
library(dlookr)
library(plyr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(ggcorrplot)
library(semTable)
library(magick)
library(mirt)
library(gtsummary)
library(gt)
library(gtExtras)
library(likert)
library(VIM) #Missing data
library(kutils)
library(simsem)
library(tabledown)
library(readxl)
library(koRpus)# stable release
library(koRpus.lang.en)
library(reshape)
library(ggsci)
r_refs("references.bib")
```
```{r Data, include=FALSE}
#This chunk holds code for data wrangling

data <- readRDS("leba_2021-09-08.rds")

# Separating EFA and CFA samples with descriptive column
descriptives.data <- data

## Merge "0"s and "1"s into "1"s select,subset assign)
descriptives.data[ , 9:56 ][ descriptives.data[ , 9:56 ] == 0 ] <- 1
#EFA.descriptives <- descriptives.data[1:428,]
EFA.descriptives <- subset(descriptives.data, IncludedInEFA == "TRUE")
#CFA.descriptives <- descriptives.data[429:690, ]
CFA.descriptives<- subset(descriptives.data, IncludedInEFA == "FALSE")

#Separating the EFA and CFA(only items)
sem.data <- data[, 9:56] #EFA & CFA data
sem.data[ sem.data == 0] <- 1 #Merged "0"s and "1"s into "1"s

## renaming the column-header
prefix <- "item"
sufix <- c(1:48)
colnam <- paste(prefix,sufix, sep = "" )
colnames(sem.data) <- colnam
names(sem.data)[names(sem.data) == "item1"] <- "item01"
names(sem.data)[names(sem.data) == "item2"] <- "item02"
names(sem.data)[names(sem.data) == "item3"] <- "item03"
names(sem.data)[names(sem.data) == "item4"] <- "item04"
names(sem.data)[names(sem.data) == "item5"] <- "item05"
names(sem.data)[names(sem.data) == "item6"] <- "item06"
names(sem.data)[names(sem.data) == "item7"] <- "item07"
names(sem.data)[names(sem.data) == "item8"] <- "item08"
names(sem.data)[names(sem.data) == "item9"] <- "item09"
```

```{r prepareDescTable, include=FALSE}
#This chunk holds code for demographic data wrangling.

listdescVars <- colnames(descriptives.data[c(1:3,5:7)])

#seperating naming and  reducing data for descriptive Table 
desctable.data <- dplyr::select(descriptives.data, c(listdescVars,
                          IncludedInCFA))

#renaming TRUE and FALSE from the IncludedInCFA to EFA and CFA for Descriptive Table
desctable.data$IncludedInCFA[desctable.data$IncludedInCFA == T ] <-
 "2. CFA Sample"
desctable.data$IncludedInCFA[desctable.data$IncludedInCFA == F ] <- 
 "1. EFA Sample"

# recode the Gender factor so it will only show "Gender Diverse" in the summary table
desctable.data$slypos_demographics_gender.factor <-
  recode_factor(desctable.data$slypos_demographics_gender.factor, "No" = "Yes",
         "Yes"= "No")

# create descriptive table for the demographic vars (excluding tz & Country) with gtsummary
desctable.data %>%
  tbl_summary(
    by = IncludedInCFA,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 2,
    label = list(slypos_demographics_age ~ "Age",
                 slypos_demographics_sex.factor ~ "Sex",
                 slypos_demographics_gender.factor ~ "Gender-Variant Identity",
                 slypos_demographics_language.factor ~ "Native English Speaker",
                 slypos_demographics_work_or_school.factor ~ "Occupational Status",
                 slypos_demographics_school.factor ~ "Occupational setting"
                 ),
    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% add_p() %>% add_q() %>%
modify_header(label ~ "**Variable**") %>% separate_p_footnotes() %>%
  modify_caption("Demographic Characteristics") -> desc_table

as_tibble(desc_table) -> desc_tibble 
as_kable_extra(desc_table, format = "latex",booktabs = T) -> desc_kable #save it as a knitr::kable

#summarise country/time zone data

descriptives.data%>%
  tbl_summary(
    label = slypos_demographics_tz.factor ~ "Time zone - Country",
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    missing = "no",
    sort = all_categorical() ~ "frequency",
    include = "slypos_demographics_tz.factor"
  )  %>% bold_labels() %>% 
  modify_header(label ~ "") %>%
  as_tibble(format = 'pipe') -> tz_tibble

Table \@ref(tab:tabtabtab)

```{r tabtabtab, warning=F, message=F}
desc_kable  %>% kable_styling(latex_options = c("scale_down")) %>% landscape()
```

Table \@ref(tab:tabtabtab)
